---
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    
params:
  nces_num: '1'
  years: 2000:2020
  state_abb: 'NC'
  grades: 6:12
  org_level: 'schools'
  
---

```{r setup, include=FALSE, results = 'hide'}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(glue)
library(forcats)
library(stringr)
library(tigris)
library(purrr)
library(dplyr)
library(highcharter)
library(schoolreportr)
library(flexdashboard)
```

```{r loadData, results = 'hide', include = FALSE}
# years data is available
saipe_years <- schoolreportr:::data_years_available('saipe')
ccd_years <- schoolreportr:::data_years_available('ccd')

state_fips <- schoolreportr:::state_fips_code(state_abb)

school_directory <- schoolreportr:::get_ccd_directory(org_level, nces_num, max(ccd_years))

# you should only have one district. Produce error if there are more than one.
if (length(unique(school_directory$leaid)) > 1) stop("All schools must be in the same school district")

district_leaid_number <- unique(school_directory$leaid)

# for school name, use the school if there is only one, otherwise use the district
if (length(nces_num) == 1) {
  school_names <- stringr::str_to_title(school_directory$school_name)
} else {
  distric_name <- unique(school_directory$lea_name) %>% str_to_title()
  if (length(distric_name) > 1) stop("There are more than one districts. You can only have one district")
  school_names <- stringr::str_to_title(paste0(distric_name," Schools"))
}

school_enrollment <- schoolreportr:::get_ccd_enrollment(org_level, nces_num, years, grades)

district_poverty <- schoolreportr:::get_district_in_poverty(district_leaid_number, max(saipe_years))

# for grades 9-12 in state assessments by school, they are lumped together as 9th grade
assessment_grades <- unique(ifelse(grades > 9, 9, grades))

state_assessment <- schoolreportr:::get_edfacts_state_assessments(org_level, state_fips, years, assessment_grades) %>%
  schoolreportr:::assessment_scores_by_race(org_level, nces_num) %>%
  mutate(geography = recode(geography, 
                           'state total' = glue::glue("{state_abb} Total"),
                            'schools / districts' = school_names))

# get district shapefile
district_shapefile <- school_districts(state = state_abb, cb = FALSE) %>%
  # filter for district
  filter(GEOID == district_leaid_number)

dashboard_title <- str_c(school_names, " Landscape Analysis")
```

```{r results = 'hide', include = FALSE}
max_edfacts_year <- max(state_assessment$year)
edfacts_year <- glue("{max_edfacts_year}-{max_edfacts_year}")

max_ccd_year <- max(school_directory$year)
ccd_year <- glue("{max_ccd_year}-{max_ccd_year}")

lowest_grade <- min(grades)
highest_grade <- max(grades)
grade_span <- glue("{lowest_grade}-{highest_grade}")
```

<script>
document.querySelector(".navbar-header > span.navbar-brand").innerHTML = "`r dashboard_title`";
</script> 


School/s Overview {data-orientation=rows}
=====================================   

Row
-----------------------------------------------------------------------

### Enrollment at `r school_names`<br>`r glue("{ccd_year}<br>Grades {grade_span}")`

```{r}
# get school enrollment in the needed grades for most recent year
latest_year_enrollment <- school_enrollment %>%
  filter(race == 'Total', year == max(years)) %>%
  pull(enrollment) %>%
  as.numeric() %>%
  scales::comma()

valueBox(latest_year_enrollment, icon = "fa-child")
```

### FTE Teachers at `r school_names`<br>`r glue("{ccd_year}<br>Grades {grade_span}")`

```{r}
teachers_elementary <- sum(school_directory$teachers_fte) %>%
  as.numeric() %>%
  scales::comma()

valueBox(teachers_elementary, icon = "fa-chalkboard-teacher")
```

### Percent of 5-17 year olds in poverty within district<br>`r max(saipe_years)`

```{r}
percent_students_poverty <-  district_poverty$est_population_5_17_pct[1] %>%
  scales::percent()

valueBox(percent_students_poverty, icon = "fa-bed")
```

### Percent passing math state assessments<br>`r glue("{edfacts_year}<br>Grades {grade_span}")`

```{r}
# get most recent year for the district of state assessments
school_assessment_recent_year <-  state_assessment %>%
  filter(year == !!max_edfacts_year, !str_detect(geography, 'Total'), race == 'Total')

math_assess <- school_assessment_recent_year %>%
  pull(math_pct_pass)

math_assess <- scales::percent(math_assess / 100, accuracy = 1)

valueBox(math_assess, icon = "fa-calculator")
```

### Percent passing reading state assessments<br>`r glue("{edfacts_year}<br>Grades {grade_span}")`

```{r}
read_assess <- school_assessment_recent_year %>%
  pull(read_pct_pass)

read_assess <- scales::percent(read_assess / 100, accuracy = 1)

valueBox(read_assess, icon = "fa-book")
```

Row
-------------------------------------

### District Map {data-width=700}

```{r include = FALSE}
# must create the map in a different code chunk than the chunk it is printed in
# this is because sf prints a statement in one of the commands and the only
# way to suppress the statement is the include the header include = FALSE
# but, this header prevents the map from showing as well
district_map <- schoolreportr:::leaflet_district_schools(district_shapefile, school_directory)
``` 

```{r}
district_map
```

### Cities in District {data-width=200}

```{r citiesTable, include = FALSE, results = 'hide'}
pop_colname <- glue("**City Pop.<br>{max(years)-1}**")

cities_table <- schoolreportr:::table_cities_in_single_district(district_shapefile, state = state_abb) %>% 
  cities_in_district(pop_colname)
```

```{r}
cities_table
```

School Demographics {data-orientation=rows}
=====================================  

Row {data-height=400}
-----------------------------------------------------------------------

### Total Enrollment by Year

```{r}
total_enrollment <- school_enrollment %>%
  filter(race == 'Total')

school_year_x_label <- 'School Year'

plt_tooltip <- paste0("<b>", school_names, ":</b>  {point.y:.0f} students")

hchart(total_enrollment, "line", hcaes(x = school_year_both, y = enrollment),
                    tooltip = list(pointFormat = plt_tooltip))  %>%
  schoolreportr:::add_title_axis_labels(
    plt_title = glue("{school_names}\nStudent Enrollment by School Year (grades {grade_span})"),
    plt_x_label = school_year_x_label,
    plt_y_label = glue("Enrollment in grades {grade_span}")
  )  %>%
    highcharter::hc_exporting(enabled = TRUE)
```

Row {data-height=400}
-----------------------------------------------------------------------

### Enrollment by Race and Year

```{r}
race_y_var_title <- glue("Percentage Enrollment by Race (grades {grade_span})")

school_enrollment %>%
  schoolreportr:::clean_enrollment_by_race() %>%
  schoolreportr:::hc_plot_grouped_line(
    x_col = 'school_year_both', y_col = 'percent_race', group_col = 'race', 
    plt_title = NULL, 
    x_var_title = school_year_x_label, y_var_title = race_y_var_title,
    y_percentage = TRUE
  ) %>%
  schoolreportr:::custom_hc_tooltip(create_html_tooltip('race', 'percent_race_clean', 'enrollment'))
```

State Assessments {data-orientation=rows}
=====================================  

Row {data-height=400}
-----------------------------------------------------------------------

```{r}
state_assessment <- state_assessment %>%
  # make sure the different district names are all the same
  mutate(
    # percentages for plotting
    read_pct_pass_clean = scales::percent(read_pct_pass / 100, accuracy = 1),
    math_pct_pass_clean = scales::percent(math_pct_pass / 100, accuracy = 1)
  )

all_years_total <- state_assessment %>%
  filter(race == 'Total')

# create plots for math and reading
subjects <- c('math', 'read')
plt_titles <- c('Percent passing math state assessments', 'Percent passing reading state assessments')

all_year_plts <- map2(subjects, plt_titles, function(.x, plt_title) {

  all_years_total %>%
    schoolreportr:::hc_plot_grouped_line(
      'school_year_both', glue('{.x}_pct_pass'), 'geography', glue("{plt_title} in grades {grade_span}"),
      NULL, 'Percent passing state assessment', y_percentage = TRUE
    ) %>%
    schoolreportr:::custom_hc_tooltip(
      schoolreportr:::create_html_tooltip('geography', glue('{.x}_pct_pass_clean'),
                                                       glue('{.x}_test_num_valid'))
    )

})

```

### Reading State Assessments - Percent Passing

```{r}
all_year_plts[[2]]
```

### Math State Assessments - Percent Passing

```{r}
all_year_plts[[1]]
```


Row {data-height=400}
-----------------------------------------------------------------------

```{r}
# get school assessment by race data
school <- state_assessment %>%
  filter(!str_detect(geography, 'Total'),
         race != 'Total')

# line plots of district assessments by race and year
all_years_race_district <- map2(subjects, plt_titles, function(.x, plt_title) {

  school %>%
    schoolreportr:::hc_plot_grouped_line(
      'school_year_both', glue('{.x}_pct_pass'), 'race', NULL,
      NULL, glue("{plt_title} in grades {grade_span}"), y_percentage = TRUE
    ) %>%
    schoolreportr:::custom_hc_tooltip(
      schoolreportr:::create_html_tooltip('race', glue('{.x}_pct_pass_clean'),
                                                  glue('{.x}_test_num_valid'))
      )

})

# bar  chart comparing most recent year with state averages
all_years_race <- state_assessment %>%
  filter(race != 'Total') %>%
  arrange(race)

all_years_race_total <- map2(subjects, plt_titles, function(.x, plt_title) {

  schoolreportr:::hc_plot_grouped_bar(all_years_race, 'race', glue('{.x}_pct_pass'), 'geography', glue("{plt_title} in grades {grade_span}"))

})
```

### School Reading State Assessments by Race - Percent Passing

```{r}
all_years_race_district[[2]]
```

### School and State Reading Assessments by Race - Percent Passing

```{r}
all_years_race_total[[2]]
```


Row {data-height=400}
-----------------------------------------------------------------------

### School Math State Assessments by Race - Percent Passing

```{r}
all_years_race_district[[1]]
```

### School and State Math State Assessments by Race - Percent Passing

```{r}
all_years_race_total[[1]]
```


Community Demographics {data-orientation=rows}
=====================================  

Row {data-height=350}
-------------------------------------

```{r results = 'hide'}
# find the census tracts in the school district
tracts_in_district <- schoolreportr:::census_tracts_in_district(state_abb, district_shapefile)
```

### Racial Composition

```{r}
race_popualtion_variables <- c(
  B03002_001 = 'Total',
  B03002_004 = 'Black / African-American', B03002_012 = 'Hispanic / Latinx',
  B03002_003 = 'White', B03002_005 = 'American Indian or Alaska Native', B03002_006 = 'Asian',
  B03002_007 = 'Native Hawaiian or other Pacific Islander', 
  B03002_010 = 'Two or more races', B03002_011 = 'Two or more races'
)
  
census_demographics <- schoolreportr:::acs_demographic_data(
    names(race_popualtion_variables), state_abb, tracts_in_district$tracts_in_district
  ) %>%
  mutate(variable = recode(variable, !!!race_popualtion_variables)) 

census_demographics_total <- census_demographics%>%
  schoolreportr:::calculate_percentages('Total') %>%
  mutate(variable = schoolreportr:::rename_reorder_race_education(variable)) %>%
  arrange(variable)

# overall racial composition
schoolreportr:::hc_plot_grouped_bar(census_demographics_total, 'variable', 'perc_demo', 'geography', "Racial breakdown of community")
```

### Percentage Black by Census Tract

```{r}
# calculate percentage pop in each race
races_to_use <- c('Black / African-American', 'Hispanic / Latinx', 'White')

# make sure total population value is present, because future operations rely on it
if (!"Total" %in% census_demographics$variable) stop("Population census tract data does not contain total populations.", call. = FALSE)

all_race_population <- acs_tracts_race(census_demographics, races_to_use)

schoolreportr:::choropath_tracts_race(all_race_population, 'Black / African-American')
```

Row {data-height=350}
-------------------------------------

### Percentage Hispanic / Latinx by Census Tract

```{r}
schoolreportr:::choropath_tracts_race(all_race_population, 'Hispanic / Latinx')
```

### Percentage White by Census Tract

```{r}
schoolreportr:::choropath_tracts_race(all_race_population, 'White')
```

Row {data-height=350}
-------------------------------------

### Educational Attainment of Population 25 and Over

```{r}
census_education <- schoolreportr:::educational_attainment(state_abb, tracts_in_district$tracts_in_district, max(years))

ed_order <- unique(census_education$variable)

census_education <- census_education %>%
  mutate(variable = factor(variable, levels = ed_order)) %>%
  arrange(variable)

schoolreportr:::hc_plot_grouped_bar(census_education, 'variable', 'perc_demo', 'geography', "Educational attainment of community")
```
