---
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    
params:
  district_lea_id: '1'
  year: 2019
  
---

```{r setup, include=FALSE, results = 'hide'}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# library(tntpr)
# library(flexdashboard)
# library(glue)
# library(here)
# library(tidycensus)
# library(leaflet)
# library(highcharter)
# library(gt)
# library(sf)
# library(scales)
# library(tidyverse)

# source(here('global_functions.R'))
# source(here('create-dashboard', 'dashboard_functions.R'))
# source(here('create-dashboard', 'dashboard-functions-census.R'))

nces_number <- '051068001566'
# north little rock

district_leaid <- params$district_lea_id # school_district_numbers()$district_lea[[1]]

years <- 2000:2019

```

```{r importDataDistrictOverview, results = 'hide'}
district_information <- read_rds(here(path_to_raw_data(), 'district_information.rds')) %>%
  filter(leaid == !!district_leaid)

school_information <- read_rds(here(path_to_raw_data(), 'school_information.rds')) %>%
  filter(leaid %in% !!district_leaid) %>%
  mutate(year = as.numeric(year))

full_school_name <- school_information$school_name

# name of HS in the district to be used in regular expressions
re_school_name <- school_information$school_name %>%
  str_to_lower() %>%
  # extract first 6 letters word
  str_extract("^.{6}")
  
ncessch_in_district <- school_information$ncessch

district_shapefile <- read_rds(here(path_to_raw_data(), 'shapefile_district.rds')) %>%
  filter(GEOID %in% !!district_leaid)

district_name <- district_information$lea_name[[1]]

dashboard_title <- str_c(full_school_name, " Landscape Analysis")
```

<script>
document.querySelector(".navbar-header > span.navbar-brand").innerHTML = "`r dashboard_title`";
</script> 

```{r pageDistrictSetup}
district_year <- as.numeric(district_information$year[[1]])
district_year <- glue("SY {district_year} - {district_year + 1}")
```

District Overview {data-orientation=rows}
=====================================   
   
Row
-----------------------------------------------------------------------

### Grades in School

```{r}
school_grades <- glue("{school_information$lowest_grade_offered} - {school_information$highest_grade_offered}")

valueBox(school_grades, icon = "fa-school")
```

### School Enrollment<br>`r district_year`

```{r}
school_enrollment <- school_information$enrollment[[1]] %>%
  as.numeric() %>%
  scales::comma()

valueBox(school_enrollment, icon = "fa-child")
```

### FTE Teachers at School<br>`r district_year`

```{r}
school_teachers <- school_information$teachers_fte[[1]] %>%
  as.numeric() %>%
  scales::comma()

valueBox(school_teachers, icon = "fa-chalkboard-teacher")
```

### Free or Reduced Lunch Students<br>`r district_year`

```{r}
free_reduced_lunch <- school_information$free_or_reduced_price_lunch[[1]] %>%
  as.numeric() %>%
  scales::comma()

valueBox(free_reduced_lunch, icon = "fa-utensils")
```

Row
-------------------------------------

### District Map {data-width=700}

```{r include = FALSE}
# must create the map in a different code chunk than the chunk it is printed in
# this is because sf prints a statement in one of the commands and the only
# way to suppress the statement is the include the header include = FALSE
# but, this header prevents the map from showing as well
district_map <- leaflet_district_schools(district_shapefile, school_information, 
                         state_fips = state_fips_number(), year = year)
``` 

```{r}
district_map
```
    
### Cities in District {data-width=200}

```{r citiesTable, include = FALSE, results = 'hide'}
pop_colname <- glue("**City Pop.<br>{year}**")

cities_table <- table_cities_in_single_district(district_shapefile, state_fips = state_fips_number(), year = year) %>%
  gt() %>%
  cols_label(
    NAME = md("**City Name**"),
    value = md(pop_colname)
  )
```

```{r}
cities_table
```

School Demographics {data-orientation=rows}
=====================================  

```{r importDataSchoolDemographics}
lep_enroll <- read_rds(here(path_to_raw_data(), 'enrollment_lep.rds')) %>%
  mutate(
    sy = glue("{year}-{year+1}"),
    perc_cleaned = scales::percent(perc_lep, accuracy = 1),
    perc_lep = perc_lep * 100
  ) %>%
  filter(str_detect(str_to_lower(school_name), !!re_school_name))

enroll_race_total <- read_rds(here(path_to_raw_data(), 'enrollment_race.rds')) %>%
  filter(str_detect(str_to_lower(school_name), !!re_school_name))

total_enroll <- enroll_race_total %>%
  filter(race == 'Total') %>%
  group_by(school_name, year) %>%
  summarize(total_students = sum(enrollment), .groups = 'drop') %>%
  mutate(sy = glue("{year}-{year+1}"))

race_enroll <- enroll_race_total %>%
  filter(race != 'Total') %>%
  group_by(school_name, year, race) %>%
  summarize(total_students = sum(enrollment), .groups = 'drop_last') %>%
  mutate(perc_by_race = total_students / sum(total_students)) %>%
  ungroup() %>%
  # required formatted columns for plotting
  mutate(
    sy = glue("{year}-{year+1}"),
    perc_cleaned = scales::percent(perc_by_race, accuracy = 1),
    perc_by_race = perc_by_race * 100
  )
  
```

Row {data-height=400}
-----------------------------------------------------------------------

### Total Enrollment by Year

```{r plotTotalEnrollment}
school_year_x_label <- 'School Year'

tooltip_html <- create_html_tooltip('school_name', '', 'total_students') %>%
  str_remove("[(]") %>%
  str_remove("[)]")

hchart(total_enroll, "line", hcaes(x = sy, y = total_students))  %>%
 custom_hc_tooltip(tooltip_html) %>%
  add_title_axis_labels(
    plt_title = glue("{full_school_name}\nStudent Enrollment by School Year"),
    plt_x_label = school_year_x_label,
    plt_y_label = glue("Enrollment")
  )
```

Row {data-height=400}
-----------------------------------------------------------------------

### Enrollment by Race

```{r plotRaceEnrollment}
tooltip_html <- create_html_tooltip('race', 'perc_cleaned', 'total_students')

race_enroll %>%
  arrange(desc(year)) %>%
  create_school_demo_linechart(
    x_col = 'sy', y_col = 'perc_by_race', group_col = 'race', 
    tool_tip_html = tooltip_html, 
    plt_title = "Enrollment by Race",
    x_var_title = school_year_x_label,
    y_var_title = "Percentage Enrollment by Race"
  )
```

### Enrollment by LEP Status

```{r plotLEPEnrollment}
tooltip_lep_html <- create_html_tooltip('school_name', 'perc_cleaned', 'lep_limited')

lep_plt_title <- glue("Enrollment by Limited English Proficient (LEP) Status")

hchart(lep_enroll, "line", hcaes(x = sy, y = perc_lep))  %>%
 custom_hc_tooltip(tooltip_lep_html) %>%
  add_title_axis_labels(
    plt_title = lep_plt_title,
    plt_x_label = school_year_x_label,
    plt_y_label = '% LEP'
  ) %>%
  hc_yAxis(
    labels = list(format = '{value}%'),
    min = 0, max = 100
  )
```

SAT / ACT Participation and Results  {data-orientation=rows}
===================================== 

```{r importDataSAT}
# all SAT and ACT data

act <- read_rds(here(path_to_raw_data(), 'sat_act.rds')) %>%
  filter(str_detect(str_to_lower(school_name), !!re_school_name)) %>%
  mutate(part_rate_text = glue("{round(part_rate, 0)}%")) %>%
  rename(demo = group)

act_total <- act %>%
  filter(demo == 'All Students')

unique_racial_categories <- c("African American", "American Indian", "Asian", "Hispanic", "Multiracial",
                              "Pacific Islander", "White")

act_race <- act %>%
  filter(demo %in% !!unique_racial_categories) %>%
  # make two race groups White and Students fo Color
  mutate(demo = recode(demo, 'Hispanic' = 'Hispanic / Latinx', 'African American' = 'Black'))

act_econ <- act %>%
  filter(str_detect(demo, '[E|e]conomically')) %>%
  mutate(demo = str_to_title(demo))

act_esl <- act %>%
  filter(str_detect(demo, "^Bil|^Not Bil"))
```

Row {data-height=400}
-----------------------------------------------------------------------

### Total SAT and ACT Participation

```{r}
sat_x_label <- 'Graduating Class'

tooltip_act_html <-  create_html_tooltip('school_name', 'part_rate_text', 'grads_mskd')

hchart(act_total, "line", hcaes(x = year, y = part_rate))  %>%
 custom_hc_tooltip(tooltip_act_html) %>%
  add_title_axis_labels(
    plt_title = glue("{full_school_name}\nPercentage of Students Taking Either SAT or ACT"),
    plt_x_label = sat_x_label,
    plt_y_label = "% Taking SAT or ACT"
  ) %>%
  hc_legend(enable = FALSE)
```

Row {data-height=400}
-----------------------------------------------------------------------

### SAT results - All Students

```{r}
sat_title <- "Average SAT Score"
act_title <- "Average ACT Composite Score"
sat_y_label <- "Average SAT Total Score"
act_y_label <- "Average ACT Composite Score"

tooltip_act_html <-  create_html_tooltip('school_name', 'sat_total', 'sat_exnees_mskd')

hchart(act_total, "line", hcaes(x = year, y = sat_total))  %>%
 custom_hc_tooltip(tooltip_act_html) %>%
  add_title_axis_labels(
    plt_title = sat_title,
    plt_x_label = sat_x_label,
    plt_y_label = sat_y_label
  )%>%
  hc_legend(enable = FALSE)
```

### ACT results - All Students

```{r}
tooltip_act_html <-  create_html_tooltip('school_name', 'act_compos', 'act_exnees_mskd')

hchart(act_total, "line", hcaes(x = year, y = act_compos))  %>%
 custom_hc_tooltip(tooltip_act_html) %>%
  add_title_axis_labels(
    plt_title = act_title,
    plt_x_label = sat_x_label,
    plt_y_label = act_y_label
  )
```

Row {data-height=400}
-----------------------------------------------------------------------

```{r satActRacePlots}
# create two plots by race - SAT and ACT
sat_act_race_args <- list(
  takers = c('sat_exnees_mskd', 'act_exnees_mskd'),
  y = c('sat_total', 'act_compos'),
  plt_title = c(glue("{sat_title} By Race"), glue("{act_title} By Race")),
  y_title = c(sat_y_label, act_y_label)
)

sat_race_plts <- pmap(sat_act_race_args, function(takers, y, plt_title, y_title) {
  
  
  tooltip_html <- create_html_tooltip('demo', y, takers)

  act_race %>%
    arrange(desc(year)) %>%
    create_school_demo_linechart(
      x_col = 'year', y_col = y, group_col = 'demo', 
      tool_tip_html = tooltip_html, 
      plt_title = plt_title,
      x_var_title = sat_x_label,
      y_var_title = y_title,
      y_percentage = FALSE
    )
  
})
```

### SAT results by race

```{r}
sat_race_plts[[1]]
```

### ACT results by race

```{r}
sat_race_plts[[2]]
```

Graduation Rates {data-orientation=rows}
===================================== 

```{r importGradRates}
# graduation rate data set does not have ncess numbers, so get school names of district
unique_schools_re <- unique(act$school_name) %>%
  # extract the first six letters of school name for matching
  str_extract("......") %>%
  # make regular expression so we can detect any of the patterns
  str_to_lower() %>%
  str_c(collapse = '|')

grad_rates <- read_rds(here(path_to_raw_data(), 'graduation_rates.rds')) %>%
  filter(str_detect(str_to_lower(school_name), unique_schools_re)) %>%
  mutate(
    total_number = replace_na(total_number, 'Unknown'),
    perc_grad_cleaned = scales::percent(grad_rate, accuracy = 1),
    grad_rate = grad_rate * 100,
    sy = glue("{year-1}-{year}"),
    sy = fct_reorder(sy, year)
  ) %>%
  arrange(year)

unique_sy <- levels(grad_rates$sy)

grad_rates_total <- grad_rates %>%
  filter(demographic_description == 'Overall',
         group == 'Total') 

grad_rates_race <- grad_rates %>%
  filter(group == 'Race')
```

Row {data-height=400}
-----------------------------------------------------------------------

### Overall Graduation Rates

```{r}
tooltip_grad_html <-  create_html_tooltip('school_name', 'perc_grad_cleaned', 'total_number')

grad_rates_total %>%
  arrange(year) %>%
  hchart("line", hcaes(x = sy, y = grad_rate))  %>%
   custom_hc_tooltip(tooltip_grad_html) %>%
    add_title_axis_labels(
      plt_title = glue("{full_school_name}\n4-Year Graduation Rate"),
      plt_x_label = school_year_x_label,
      plt_y_label = "4-Year Graduation Rate"
    ) %>%
    hc_legend(enabled = FALSE)
```

Row {data-height=400}
-----------------------------------------------------------------------

### Graduation Rates by Race

```{r}
tooltip_grad_race_html <-  create_html_tooltip('demographic_description', 'perc_grad_cleaned', 'total_number')

grad_rates_race %>%
  create_school_demo_linechart(
    x_col = 'sy', y_col = 'grad_rate', group_col = 'demographic_description', 
    tool_tip_html = tooltip_grad_race_html, 
    plt_title = '4-Year Graduation Rate by Race',
    x_var_title = school_year_x_label,
    y_var_title = "4-Year Graduation Rate",
    y_percentage = TRUE, reverse_x = FALSE
  ) %>%
  hc_xAxis(categories = unique_sy)
```

College Enrollment {data-orientation=rows}
===================================== 

```{r}
higher_ed <- read_rds(here(path_to_raw_data(), 'higher_ed_enrolled.rds')) %>%
  filter(str_detect(str_to_lower(school_name), unique_schools_re)) %>%
  mutate(
    perc_cleaned = scales::percent(perc_total, accuracy = 1),
    perc_total = perc_total * 100
  )
```

Row {data-height=400}
-----------------------------------------------------------------------

### College Enrollment in a TX Public Institution

```{r}
tooltip_higher_ed_html <-  create_html_tooltip('institution_name', 'perc_cleaned', 'num_students')

higher_ed %>%
  create_school_demo_linechart(
    x_col = 'year', y_col = 'perc_total', group_col = 'institution_name', 
    tool_tip_html = tooltip_higher_ed_html, 
    plt_title = glue("{full_school_name}\nPercentage of Students Enrolled in Higher Education at\n
                      TX Public Institution"),
    x_var_title = school_year_x_label,
    y_var_title = "4-Year Graduation Rate",
    y_percentage = TRUE, reverse_x = TRUE
  )
```

Community Demographics {data-orientation=rows}
===================================== 

```{r countiesInDistrict}
# texas block groups
tx_census_tracts_shapefile <- tigris::tracts(
  state = state_fips_number(),
  year = year
)  %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(4326, quiet = TRUE)

district_shapefile <- district_shapefile %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(4326, quiet = TRUE)

# extract tracts that touch the district
tracts_in_district <- district_blockgroups(district_shapefile, tx_census_tracts_shapefile)

tracts_geoid_in_district <- tracts_in_district$GEOID

# only keep tracts within the district for county shappefile
tx_census_tracts_shapefile <- tx_census_tracts_shapefile %>%
  filter(GEOID %in% !!tracts_geoid_in_district)

# centroid of district, used to center view of map
district_centroid <- st_centroid(district_shapefile$geometry) %>%
  map_dfr(~tibble(lng = .x[[1]], lat = .x[[2]]))
```

Row {data-height=350}
-------------------------------------

District demographic information reflects the demographic data from all census tracts that have any 
area within the school district.
The map below shows these census tracts. The dark colored shape represents the boundaries for `r district_name`.
The light colored shapes are the census tracts used for calculating demographic data.

### Census Tracts within District

```{r}
district_map %>%
  addPolygons(
      data = tx_census_tracts_shapefile$geometry,
      color = "#444444", weight = 1, smoothFactor = 0.5,
      opacity = 1.0, fillOpacity = .1,
      layerId = 1
    )
```

Row {data-height=350}
-------------------------------------

```{r racePopulationData}
acs_vars <- tidycensus::load_variables(max(years), "acs5", cache = TRUE)

population_tables <- c(
  Total = 'B03002_001',
  White = 'B03002_003',
  Black = 'B03002_004',
  Latinx = 'B03002_012'
)

race_percentages <- acs_district_state(population_tables, state_fips_number(), tracts_geoid_in_district)  %>%
  calculate_percentages('Total')
```
  
### Racial Composition

```{r}
perc_tooltip <- list(valueDecimals = 0, valueSuffix = "%", shared = TRUE)

plt_hc_barchart_percent(race_percentages, perc_tooltip) %>%
  plt_hc_percentage('Racial Composition')
```

### Educational Attainment of Population 25 and Over

```{r eduacationAttainmentData}
# list of acs variables for educational attainment
education_variables <- str_pad(1:25, width = 3, side = 'left', pad = '0')
education_variables <- str_c('B15003_', education_variables)

# get county and state educational attainment numbers
education <- acs_district_state(education_variables, state_fips_number(), tracts_geoid_in_district) %>%
  left_join(acs_vars[c('name', 'label')], by = c('variable' = 'name'))

# re-bin educational attainment levels
education <- education %>%
  # use variable numbers to bin groups
  mutate(variable_number = str_extract(variable, "[0-9]{3}$") %>% as.numeric()) %>%
  mutate(education_level = case_when(
    variable_number == 1 ~ 'Population 25 and over',
    between(variable_number, 2, 16) ~ 'No HS diploma or GED',
    between(variable_number, 17, 18) ~ 'HS diploma or GED',
    between(variable_number, 19, 21) ~ "Some college or associates degree",
    between(variable_number, 22, 22) ~ "Bachelor's degree",
    between(variable_number, 23, 25) ~ "Beyond bachelor's",
    TRUE ~ 'Failed to match'
  )) %>%
  # create ordered factor based on variable number
  group_by(education_level) %>%
  mutate(max_variable_number = max(variable_number)) %>%
  ungroup() %>%
  mutate(education_level = fct_reorder(education_level, max_variable_number)) %>%
  # don't need these variables because we have descriptive labels for them
  select(-variable_number, -variable, -label) %>%
  #rename tpo match function that calculates percentages
  rename(variable = education_level)

# calculate percentages for each newly created bin
education_percentages <- education %>%
  calculate_percentages('Population 25 and over')
```

```{r pltEducationalAttainment}
plt_hc_barchart_percent(education_percentages, perc_tooltip) %>%
    plt_hc_percentage('Percentage of Population 25 and Over')
```
