---
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    
params:
  district_lea_id: '1'
  year: 2019
  
---

```{r setup, include=FALSE, results = 'hide'}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

district_leaid_number <- '200001' # district in alaska
state_abb <- 'AK'
years <- 2016:2017

devtools::load_all()
```

```{r loadData, results = 'hide', include = FALSE}
# pull state assessment data from API
state_assessment <- get_state_assessments_by_district(state_fips_code(state_abb), years)

```

```{r loadData, results = 'hide', include = FALSE}

data_sources_to_import <- c('enrollment race', 'enrollment lep', 'state assessments')

# get all Urban Institute Education Data Portal API data
ed_data <- get_all_educationdata(district_leaid_number, years, data_sources_to_import)

school_directory <- ed_data$directory$data %>%
  dplyr::mutate(dplyr::across(school_name:city_location, ~stringr::str_to_title(.x)))

# get state to extract district and census tract shape files
state_abb <- school_directory$state_location

# get district shapefile
district_shapefile <- tigris::school_districts(state = state_abb, cb = FALSE) %>%
  # filter for district
  dplyr::filter(GEOID == school_directory$leaid)

school_name <- school_directory$school_name[1]

dashboard_title <- stringr::str_c(school_name, " Landscape Analysis")

yr <- school_directory$year[[1]]

district_year <- glue::glue("{yr} - {yr + 1}")

race_remove <- c('Total', 'Unknown')
```

<script>
document.querySelector(".navbar-header > span.navbar-brand").innerHTML = "`r dashboard_title`";
</script> 


District Overview {data-orientation=rows}
=====================================   

Row
-----------------------------------------------------------------------

### District Enrollment<br>`r district_year`

```{r}
school_enrollment <- state_assessment %>%
  as.numeric() %>%
  scales::comma()

flexdashboard::valueBox(school_enrollment, icon = "fa-child")
```

### FTE Teachers at School<br>`r district_year`

```{r}
school_teachers <- school_directory$teachers_fte[[1]] %>%
  as.numeric() %>%
  scales::comma()

flexdashboard::valueBox(school_teachers, icon = "fa-chalkboard-teacher")
```

### Free or Reduced Lunch Students<br>`r district_year`

```{r}
free_reduced_lunch <- school_directory$free_or_reduced_price_lunch[[1]] %>%
  as.numeric() %>%
  scales::comma()

perc_free_reduced <- scales::percent(school_directory$perc_free_reduced_lunch[[1]])

free_reduced_lunch <- glue::glue("{free_reduced_lunch} ({perc_free_reduced})")

flexdashboard::valueBox(free_reduced_lunch, icon = "fa-utensils")
```

Row
-------------------------------------

### District Map {data-width=700}

```{r include = FALSE}
# must create the map in a different code chunk than the chunk it is printed in
# this is because sf prints a statement in one of the commands and the only
# way to suppress the statement is the include the header include = FALSE
# but, this header prevents the map from showing as well
district_map <- leaflet_district_schools(district_shapefile, school_directory)
``` 

```{r}
district_map
```

### Cities in District {data-width=200}

```{r citiesTable, include = FALSE, results = 'hide'}
pop_colname <- glue::glue("**City Pop.<br>{max(years)}**")

cities_table <- table_cities_in_single_district(district_shapefile, state = state_abb, year = max(years)) %>%
  cities_in_district()
```

```{r}
cities_table
```

School Demographics {data-orientation=rows}
=====================================  

Row {data-height=400}
-----------------------------------------------------------------------

### Total Enrollment by Year

```{r}
total_enrollment <- ed_data$enrollment_race$data %>%
  dplyr::filter(race == 'Total')

school_year_x_label <- 'School Year'

plt_tooltip <- paste0("<b>", school_name, ":</b>  {point.y:.0f} students")

highcharter::hchart(total_enrollment, "line", highcharter::hcaes(x = year, y = total_enrollment),
                    tooltip = list(pointFormat = plt_tooltip))  %>%
  add_title_axis_labels(
    plt_title = glue::glue("{school_name}\nStudent Enrollment by School Year"),
    plt_x_label = school_year_x_label,
    plt_y_label = "Enrollment"
  )
```

Row {data-height=400}
-----------------------------------------------------------------------

### Enrollment by Race and Year

```{r}
race_y_var_title <- "Percentage Enrollment by Race"

race_enrollment <- ed_data$enrollment_race$data %>%
  dplyr::filter(!race %in% !!race_remove)

race_enrollment$race <- relevel_race(race_enrollment$race)

tooltip_html <- create_html_tooltip('race', '.perc_cleaned', 'total_enrollment')

race_enrollment %>%
  dplyr::arrange(year, race) %>%
  pretty_percent_cols('perc_enrollment') %>%
  hc_plot_grouped_line(
    x_col = 'year', y_col = '.perc_plot', group_col = 'race', 
    tool_tip_html = tooltip_html, 
    plt_title = NULL, 
    x_var_title = school_year_x_label, y_var_title = race_y_var_title,
    y_percentage = TRUE
  ) 
```

### Enrollment by Race with State Averages

```{r}
state_enrollment_race <- enrollment_race %>%
  dplyr::filter(
    state_fips == state_fips_code(state_abb),
    !race %in% race_remove
  ) %>%
  dplyr::mutate(location = glue::glue("{state_abb} State-wide Percentage")) %>%
  dplyr::select(location, race, perc_enrollment = perc_enrollment_state)

state_value <- unique(state_enrollment_race$location)

race_enrollment_school_state <- race_enrollment %>%
  dplyr::filter(year == max(year)) %>%
  dplyr::mutate(location = !!school_name) %>%
  dplyr::select(location, race, perc_enrollment) %>%
  dplyr::bind_rows(state_enrollment_race) %>%
  dplyr::mutate(
    perc_enrollment = perc_enrollment * 100,
    location = forcats::fct_relevel(location, state_value, after = Inf)
  )

race_enrollment_school_state$race <- relevel_race(race_enrollment_school_state$race)

race_order <- levels(race_enrollment_school_state$race)

hc_plot_grouped_bar(race_enrollment_school_state, 'race', 'perc_enrollment', 'location', 
                    race_order, race_y_var_title)
```

Row {data-height=400}
-----------------------------------------------------------------------

### Enrollment by LEP and Year

```{r}
lep_y_var_title <- "Percentage Enrollment by LEP"

lep_enrollment <- ed_data$enrollment_lep$data %>%
  dplyr::mutate(
    perc_lep = perc_lep * 100,
    perc_lep_cleaned = glue::glue("{round(perc_lep, 0)}%"),
    year = as.character(year)
  )

tooltip_html <- create_html_tooltip('year', 'perc_lep_cleaned', 'lep_limited')

highcharter::hchart(lep_enrollment, "line", highcharter::hcaes(x = year, y = perc_lep))  %>%
  custom_hc_tooltip(tooltip_html) %>%
  add_title_axis_labels(
    plt_title = NULL,
    plt_x_label = school_year_x_label,
    plt_y_label = "Enrollment"
  ) %>%
  plt_hc_percentage(lep_y_var_title)
```

### Enrollment by LEP with State Averages

```{r}
state_enrollment_lep <- state_lep %>%
  dplyr::filter(state_fips == state_fips_code(state_abb)) %>%
  dplyr::mutate(
    location = glue::glue("{state_abb} State-wide Percentage"),
    perc_lep_state = perc_lep_state * 100
  ) %>%
  dplyr::select(location, perc_lep = perc_lep_state)

lep_enrollment_school_state <- lep_enrollment %>%
  dplyr::filter(year == max(year)) %>%
  dplyr::mutate(location = !!school_name) %>%
  dplyr::select(location, perc_lep) %>%
  dplyr::bind_rows(state_enrollment_lep)

highcharter::hchart(
  lep_enrollment_school_state, "column", 
  highcharter::hcaes(x = location, y = perc_lep),
  tooltip = list(pointFormat = "<b>Percent LEP:</b> {point.y:,.0f}%")
) %>%
  plt_hc_percentage(lep_y_var_title) %>%
  highcharter::hc_xAxis(title = NULL)
```

State Assessments {data-orientation=rows}
=====================================  

Row {data-height=400}
-----------------------------------------------------------------------

```{r}
reading_y_var_title <- "Percentage Passing State Reading Assessments"

single_state_assessment <- state_assessments %>%
  dplyr::filter(state_fips == state_fips_code(state_abb))  %>%
  dplyr::mutate(location = glue::glue("{state_abb} State-wide Percentage"))

assessment <- ed_data$state_assessments$data %>%
  dplyr::filter(year == max(year))  %>%
  dplyr::mutate(location = !!school_name) %>%
  dplyr::bind_rows(single_state_assessment) %>%
  dplyr::mutate(across(contains("_pct_"), ~.x * 100)) %>%
  dplyr::mutate(across(contains("_pct_"), ~scales::percent(.x/100, accuracy = 1), .names = "{.col}_clean"))

assessment_total <- assessment %>%
  dplyr::filter(race == 'Total')

assessment_race <- assessment %>%
  dplyr::filter(!race %in% !!race_remove)
  
assessment_race$race <- relevel_race(assessment_race$race)
```

### Reading State Assessments

```{r}
hc_plot_assessments(assessment_total, 'location', 'read_test_pct_prof_midpt', "Reading")
```

### Reading State Assessments by Race

```{r}
hc_plot_grouped_bar(assessment_race, 'race', 'read_test_pct_prof_midpt', 'location', race_order, race_y_var_title)
```

Row {data-height=400}
-----------------------------------------------------------------------

### Math State Assessments

```{r}
hc_plot_assessments(assessment_total, 'location', 'math_test_pct_prof_midpt', "Math")
```

### Math State Assessments by Race

```{r}
hc_plot_grouped_bar(assessment_race, 'race', 'math_test_pct_prof_midpt', 'location', race_order, race_y_var_title)
```

Community Demographics {data-orientation=rows}
=====================================  

```{r}
# find the census tracts in the school district
tracts_in_district <- census_tracts_in_district(state_abb, max(years), district_shapefile)
```

Row {data-height=350}
-------------------------------------

### Racial Composition

```{r}
race_popualtion_variables <- c(
  B03002_001 = 'Total',
  B03002_004 = 'Black, African-American', B03002_012 = 'Hispanic / Latinx',
  B03002_003 = 'White', B03002_005 = 'American Indian or Alaska Native', B03002_006 = 'Asian',
  B03002_007 = 'Native Hawaiian or other Pacific Islander', 
  B03002_010 = 'Two or more races', B03002_011 = 'Two or more races'
)
  
census_demographics <- acs_demographic_data(names(race_popualtion_variables), state_abb, tracts_in_district$tracts_in_district) %>%
  dplyr::mutate(variable = dplyr::recode(variable, !!!race_popualtion_variables)) %>%
  calculate_percentages('Total')

hc_plot_grouped_bar(census_demographics, 'variable', 'perc_demo', 'geography', race_order, "Racial breakdown of community")
```

### Educational Attainment of Population 25 and Over

```{r}
census_education <- educational_attainment(state_abb, tracts_in_district$tracts_in_district, max(years))

ed_order <- unique(census_education$variable)

hc_plot_grouped_bar(census_education, 'variable', 'perc_demo', 'geography', ed_order, "Educational attainment of community")
```

Row {data-height=400}
-----------------------------------------------------------------------

### Racial Composition of Community by Census Tract

```{r}

```
